---
title: "Paid Parental Leave, Home Resources, and Grade 4 Mathematics"
subtitle: "Reproducibility analysis for TIMSS 2019"
author: "Armel Landry Batchi-Bouyou, Päivi Häkkinen, and Ronny Scherer"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    number_sections: true
    fig_width: 10
    fig_height: 7
---

```{r setup, include=FALSE}
project_root <- normalizePath("..", winslash = "/", mustWork = TRUE)
knitr::opts_knit$set(root.dir = project_root)
setwd(project_root)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  dpi = 150,
  cache = FALSE
)

source("R/00_setup.R")
source("R/01_helpers.R")
source("R/02_external_data.R")
source("R/03_timss_data.R")
source("R/04_descriptives.R")
source("R/05_country_gradients.R")
source("R/06_meta_regression.R")
source("R/07_supplementary.R")
```

# Scope and reproducibility note

This code-only workflow reproduces the tables and figures for the analysis of
TIMSS 2019 Grade 4 mathematics, home resources for learning, and paid parental
leave across 39 education systems. TIMSS microdata are not included. The user
must obtain the official files and set `TIMSS_DIR` in a local `.Renviron` file.

The released analysis estimates weighted linear regressions separately for the
five mathematics plausible values, uses the TIMSS final student weight
`TOTWGT`, and pools coefficients and model-based weighted-regression standard
errors with Rubin's rules. It does **not** use TIMSS replicate weights. The
manuscript must use the same description.

The analytical sequence is:

1. Descriptive system profiles and the leave-group heatmap.
2. System-specific home-resource gradients under main and reduced adjustment.
3. Random-effects meta-regression of the system gradients on paid leave,
   adjusted for GDP per capita and female labour force participation.
4. Supplementary student-level and system-level correlations.

# Data access and preparation

## External policy and macroeconomic indicators

The paid-leave crosswalk is frozen in `R/02_external_data.R` so that source
revisions do not silently change the archived results. GDP per capita and female
labour force participation are averaged across 2018 to 2020 using the World Bank
API, matching the released code.

```{r external-data}
country_external <- build_country_external()
country_external
```

## TIMSS 2019 Grade 4

```{r timss-data}
timss_2019 <- read_timss_2019()
raw_data <- extract_timss_variables(timss_2019)
analysis_data <- build_analysis_data(raw_data, country_external)
validate_analysis_data(analysis_data)

cat("Students:", format(nrow(analysis_data), big.mark = ","), "\n")
cat("Education systems:", n_distinct(analysis_data$iso3c), "\n")
```

# RQ1: Descriptive system profiles

```{r table1}
country_profiles <- build_country_profiles(analysis_data)
table1 <- make_table1(country_profiles)
table1
```

```{r figure1, fig.width=9, fig.height=5}
heatmap_data <- build_heatmap_data(analysis_data)
figure1 <- make_figure1(heatmap_data)
figure1
```

# RQ2: System-specific home-resource gradients

The main specification adjusts for parental education, preschool years, and
gender. The reduced specification omits parental education but retains the same
complete-case sample to isolate the change in adjustment.

```{r country-slopes}
country_slopes_main <- build_country_slopes(analysis_data, adjustment = "main")
country_slopes_reduced <- build_country_slopes(analysis_data, adjustment = "reduced")

cat("Main-model systems:", nrow(country_slopes_main), "\n")
cat(
  "Statistically significant main-model slopes:",
  sum(country_slopes_main$p < 0.05),
  "of",
  nrow(country_slopes_main),
  "\n"
)
```

```{r table2}
table2 <- make_slope_table(country_slopes_main, adjustment = "main")
table2
```

```{r table-s1}
table_s1 <- make_slope_table(country_slopes_reduced, adjustment = "reduced")
table_s1
```

```{r figure2, fig.width=10, fig.height=6.5}
prediction_grid <- build_prediction_grid(analysis_data)
figure2 <- make_figure2(prediction_grid)
figure2
```

```{r figure-s2, fig.width=8.5, fig.height=10}
figure_s2 <- make_figure_s2(country_slopes_main)
figure_s2
```

# RQ3: Cross-national meta-regression

The meta-regression is the only inferential test of whether leave duration is
associated with variation in the home-resource gradient across systems.

```{r meta-model}
meta_data <- build_meta_data(country_slopes_main, country_external)
meta_model <- fit_meta_regression(meta_data)
summary(meta_model)
```

```{r table3}
table3 <- make_table3(meta_model, meta_data)
table3
```

```{r figure3, fig.width=10, fig.height=7}
figure3 <- make_figure3(meta_data)
figure3
```

# Supplementary correlations

```{r table-s2}
student_correlations <- build_student_correlation_matrix(analysis_data)
table_s2 <- make_table_s2(student_correlations)
table_s2
```

```{r figure-s1, fig.width=9, fig.height=7.5}
figure_s1 <- make_figure_s1(country_profiles, country_slopes_main)
figure_s1
```

# Key numerical checks

```{r numerical-checks}
pooled_model <- metafor::rma(
  yi = resource_slope,
  vi = se^2,
  data = country_slopes_main,
  method = "REML"
)

leave_row <- broom::tidy(meta_model) %>%
  filter(term == "leave_weeks")

checks <- tibble(
  Quantity = c(
    "Education systems in slope analysis",
    "Significant system slopes",
    "Smallest system slope",
    "Largest system slope",
    "Random-effects pooled slope",
    "Leave meta-regression coefficient",
    "Leave meta-regression p value"
  ),
  Value = c(
    nrow(country_slopes_main),
    sum(country_slopes_main$p < 0.05),
    min(country_slopes_main$resource_slope),
    max(country_slopes_main$resource_slope),
    as.numeric(pooled_model$b),
    leave_row$estimate,
    leave_row$p.value
  )
)

checks
```

# Session information

```{r session-information}
sessionInfo()
```
